---
title: "Project_606"
author: "SantoshCheruku"
date: "March 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```


### Research question 
**Does the choice of STEM major in graduation have an effect on the graduate salary?**



### Cases 
**The recent grad list compiled by census.gov has provided us with a list of compiled data of grads during 2010-2012. There are observations made against 173 majors.Each observation represents a major with data mentioning the number of grads who passed, their employement rate, salary ranges etc **



### Data collection 
**The data has been collected and compiled by American Community Survey (ACS) and made available as Public Use Micro Service (PUMS) data.**



### Type of study 
**This data has been gathered as an observation study**



### Data Source 
**This data has been gathered and compiled by ACS, and made available at below locations.**
**https://github.com/fivethirtyeight/data/tree/master/college-majors**
**https://www.census.gov/programs-surveys/acs/data/pums.html**




### Dependent Variable
**Median salary of a student is the dependent or response variable and is numerical (quantitative)**



### Independent Variable
**The course major is an independent variable (qualitative) and number of grad students is another independent variable (quantitative)**



### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}
grads <- read.csv(file="C://CUNY//GIT//CUNY//Semester1//606//Project//grad-students.csv")

summary(grads$Grad_median)

summary(grads$Grad_total)

ggplot(grads, aes(x=grads$Major_category, y = Grad_median)) + geom_bar(stat="identity") + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```