---
title: "Airlines_arrivalDelay_Analysis"
author: "SantoshCheruku"
date: "March 2, 2019"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyr", repos = "http://cran.us.r-project.org" )
#install.packages("dplyr", repos = "http://cran.us.r-project.org")
#install.packages("dplyr", repos = "http://cran.us.r-project.org")

#devtools::install_github("hadley/tidyr")

library(tidyr)
library(dplyr)
library(ggplot2)
```

##Retrieve the data from a CSV file
```{r}
fullData <- read.csv2(file = "C://CUNY//GIT//CUNY//Semester1//607//Week4//data.csv", header = T, sep = ",",  na.strings = c("","NA"), blank.lines.skip = TRUE, stringsAsFactors = F)
names(fullData)[1] <- 'Airline'
```

##Convert the data from 'Wide' into a 'Long' format for ease of analysis
```{r}
fullData_Long <- gather(fullData, key="City", value="FlightCount", 3:7)
head(fullData_Long)

```

##Transform the data using tidyr and dplyr functions
    1. gather -> Convert data from Wide format to Long format 
    2. select -> Select only a limited set of columns
    3. group_by -> Group the data in a tbl/df mentioning selected columns
    4. summarize -> Generate sumarized stats
    5. filter -> Filter the tbl/df by applying conditions on columns
    6. mutate -> Modify the column data or create new columns
    7. left_join  -> Apply left join on different datasets
    
    
```{r}
airline_city_total <- fullData_Long %>% select(Airline, City, FlightCount) %>% group_by(Airline, City) %>% summarize(totalcount=sum(FlightCount))

fullData_long_totalcount <- left_join(fullData_Long, airline_city_total, by=c("Airline","City"))
head(fullData_long_totalcount)

city_airline_ontimeRatio <- fullData_long_totalcount %>% select(Airline, Status, City, FlightCount, totalcount) %>% mutate(ratio=FlightCount/totalcount) %>% filter(Status=='on time') %>% group_by(City, Airline) %>% select(City, Airline, ratio)
city_airline_ontimeRatio

final_transformed_data <- mutate(city_airline_ontimeRatio, ontime_ratio=ratio, delay_ratio=1-ratio) %>% select(City, Airline, ontime_ratio, delay_ratio)
final_transformed_data

```

##<u>**Number of operating flights**</u> by City and Airline
```{r}
ggplot(fullData_long_totalcount, aes(x=City, y = FlightCount, fill=Airline)) + geom_bar(stat="identity", position = position_dodge())

```

##Ratio of flights coming <u>**on time**</u> by City and Airline
```{r}

ggplot(final_transformed_data, aes(x=City, y = ontime_ratio, fill=Airline)) + geom_bar(stat="identity", position = position_dodge())
```

##Ratio of flights <u>**delayed**</u> by City and Airline
```{r}

ggplot(final_transformed_data, aes(x=City, y = delay_ratio, fill=Airline)) + geom_bar(stat="identity", position = position_dodge())


```

