---
title: "Data 607 Team Sparc"
author: "Priya Shaji"
date: "5/4/2019"
output: html_document
---


---
title: "Data 607 Final Project"
author: "Priya Shaji"
date: "5/3/2019"
output: html_document
---

## Load the Data

```{r eval=FALSE}
library(tidyverse)
library(psych)
library(tidyverse)
library(rpart)
library(rpart.plot)
library(party)
library(kableExtra)
```



```{r} 
library(tidyverse)
train <- read_tsv("drugLibTrain_raw.tsv", col_names = TRUE, cols(
  X1 = col_double(),
  urlDrugName = col_character(),
  rating = col_double(),
  effectiveness = col_factor(),
  sideEffects = col_factor(),
  condition = col_factor(),
  benefitsReview = col_character(),
  sideEffectsReview = col_character(),
  commentsReview = col_character()
))

test <- read_tsv("drugLibTest_raw.tsv", col_names = TRUE, cols(
  X1 = col_double(),
  urlDrugName = col_character(),
  rating = col_double(),
  effectiveness = col_factor(),
  sideEffects = col_factor(),
  condition = col_factor(),
  benefitsReview = col_character(),
  sideEffectsReview = col_character(),
  commentsReview = col_character()
))

dataset <- full_join(train, test)

```

## Analysis{.tabset}


```{r}
library(FSA)

Drugs_summary<-Summarize(dataset$rating ~ dataset$urlDrugName,
          , 
          digits=3)

head(Drugs_summary)
```



```{r}
qqnorm(dataset$rating)
qqline(dataset$rating)
```


From the above plots it is clear that the data distribution is not normal


### Statistical inference



<strong> ANOVA </strong> 

We will conduct an ANOVA analysis to test the hypothesis of, if the means of the ratings of drugs are same overall.


H0= mean rating is same for all the effective types

H1= not all means are equal


```{r}
drug.aov<-aov(dataset$rating~dataset$effectiveness )
```

```{r}
attributes(drug.aov)

summary(drug.aov)

```

Since. the p-value is less than 0.05,  we will reject the null hypothesis and conclude that not all means are equal




```{r}
#conduct a Turkey's multiple comparison procedure

#TukeyHSD(drug.aov,console=TRUE)
```


<strong>Kruskal Wallis One way analysis of variance</strong>

It is a non-parametric equivalent to the one way analysis of variance


H0= An extremely effective drug means a high rated drug

H1= An extremely effective drug is not necessarily a high rated drug

```{r}
kruskal.test(dataset$rating~dataset$effectiveness)
```


Since ,p-value is less than 0.05, we reject the null hypothesis.
Therefore an extremely effective drug is not necessarily a high rated drug.


Create a new variable which is the rating values as an ordered factor

```{r}
 

dataset$rating.f = factor(dataset$rating,
                          ordered=TRUE)

```


Summarize data treating rating  as factors

```{r}
#xtabs( ~ urlDrugName + rating.f,
 #     data = dataset)
```


Proportional table for rating of all drugs 

```{r}
#XT = xtabs( ~ urlDrugName + rating.f,
 #     data = dataset)


#prop.table(XT, 
#           margin = 1)
```



Bar plots by group
Note that the bar plots don’t show the effect of the blocking variable.

It shows the percentage of ratings for all the drugs in the dataset

```{r}
library(lattice)

par(mfrow=c(10,10))

drug_rating_plot<-histogram(~ rating.f | urlDrugName,
          data=dataset,
          layout=c(1,5)      #  columns and rows of individual plots
          )

head(drug_rating_plot)
```



## DataBase

Process

In Mysql

1) create database Data607FinalProject
2) set it as default schema
3) import tables from the local file via Table Data Import Wizard

In R

1) Connect the Mysql database using dbConnect
2) a) Case 1: if table exists, drop it
   b) Case 2: Else create a new table

3) Specify the column names and their definition
4) Write Query to update the tables
5) Close Connection





```{r}

CheckDB <- function() {
  
#--------------- Open Connection 
  
  con2 <- dbConnect(RMySQL::MySQL(), dbname = "Data607FinalProject", host="localhost", user="root",password= "Sparc@DATA_607")

#--------------- Drop Table if Exists 
  
  
if(dbExistsTable(con2, "dataset")){
    dbRemoveTable(con2,"dataset")
  
  }
 

  
  
  
#-------------- Create first table for dataset
  
  
query <- "CREATE TABLE `dataset` (  
`Drug_id`  INT AUTO_INCREMENT,
`urlDrugName` VARCHAR(45) NULL,
`rating` VARCHAR(45) NULL,
`effectiveness` VARCHAR(45) NULL, 
`sideEffects` VARCHAR(45) NULL, 
`condition` VARCHAR(45) NULL,
`benefitsReview` VARCHAR(45) NULL,
`sideEffectsReview` VARCHAR(45) NULL,
`commentsReview` VARCHAR(45) NULL,
PRIMARY KEY (`Drug_id`))"



tab1 <- dbSendQuery(con2,query)
tab1data <- dbFetch(tab1)



#--------------- Close Connection


dbDisconnect(con2)

}

#--------------- Update tables


updateDB <- function(urlDrugName, rating, effectiveness, sideEffects, condition, benefitsReview, sideEffectsReview, commentsReview){

  con2 <- dbConnect(RMySQL::MySQL(), dbname = "Data607FinalProject", host="localhost", user="root",password= "Sparc@DATA_607")
  
 
#--------------- Update dataset table 
  
  
query <- sprintf("insert into dataset (DrugName, Rating, Effectiveness, SideEffects, Condition, BenefitsReview, SideEffectsReview, CommentsReview) values('%s','%s','%s','%s','%s','%s','%s','%s')",urlDrugName, rating, effectiveness, sideEffects, condition, benefitsReview, sideEffectsReview, commentsReview)
  print (query)
tab1 <- dbSendQuery(con2,query)
tab1data <- dbFetch(tab1)

#--------------- Get 'Drug_id' to update dataset table


query <- sprintf("select Drug_id from dataset where DrugName='%s', Rating='%s', Effectiveness='%s', SideEffects='%s', Condition='%s', BenefitsReview='%s', SideEffectsReview='%s', CommentsReview='%s' limit 1",urlDrugName, rating, effectiveness, sideEffects, condition, benefitsReview, sideEffectsReview, commentsReview)
  
tab1 <- dbSendQuery(con2,query)
tab1data <- dbFetch(tab1)
data_id <- tab1data$Drug_id



#---------------- Close Connection


dbDisconnect(con2)



}







```








