---
title: "606 Week1 Homework"
author: "SantoshCheruku"
date: "February 9, 2019"
output: html_document
---

```{r include=FALSE}
library(openintro)
data(heartTr)
```

### Question: 1.8  
"Smoking habits of UK residents. A survey was conducted to study the smoking habit of UK residents. Below is a data matrix displaying a portion of the data collected in this survey.Note that "£" stands for British Pounds Sterling, "cig" stands for cigarettes, and "N/A" refers to a missing component of the data." 

*  What does each row of the data matrix represent?
   *Each row represents an individual survey response*
*  How many participants were included in the survey?
```{r}
nrow(heartTr)
```

*  Indicate whether each variable in the study is numerical or categorical. If numerical, identify as continuous or discrete. If categorical, indicate if the variable is ordinal.
  1. sex - Categorical
  2. age - Numerical (discrete)
  3. marital - Categorical
  4. grossIncome - Categorical (Ordinal)
  5. smoke - Categorical
  6. amtWeekends - Numerical(Discrete)
  7. amtWeekdays - Numerical(Discrete)


### Question 1.10
"Cheaters, scope of inference. Exercise 1.5 introduces a study where researchers studying
the relationship between honesty, age, and self-control conducted an experiment on 160 children
between the ages of 5 and 15. The researchers asked each child to toss a fair coin in private and to
record the outcome (white or black) on a paper sheet, and said they would only reward children
who report white. Half the students were explicitly told not to cheat and the others were not given
any explicit instructions. Di???erences were observed in the cheating rates in the instruction and no
instruction groups, as well as some di???erences across children's characteristics within each group."


Answers:
* Identify the population of interest and sample of study
  __*The population of interest is children in the ages of 5 and 15. The sample of study is 160 children.*__
*  Comment on whether or not the results of the study can be generalized to the population, and
if the findings of the study can be used to establish causal relationships.
  __*The findings can be generalized since it was an experiment. Yes, a causal relationship can be established.*__
  
  
### Question  1.28 Reading the paper. Below are excerpts from two articles published in the NY Times:
    (a) An article titled Risks: Smokers Found More Prone to Dementia states the following:61
    "Researchers analyzed data from 23,123 health plan members who participated in a voluntary
    exam and health behavior survey from 1978 to 1985, when they were 50-60 years old. 23 years
    later, about 25% of the group had dementia, including 1,136 with Alzheimer's disease and 416
    with vascular dementia. After adjusting for other factors, the researchers concluded that pack-aday
    smokers were 37% more likely than nonsmokers to develop dementia, and the risks went up
    with increased smoking; 44% for one to two packs a day; and twice the risk for more than two
    packs."
    Based on this study, can we conclude that smoking causes dementia later in life? Explain your
    reasoning.
__*Answer: No, we cannot conclude that smoking causes dementia since this was not a controller experiment. People were only surveyed and their responses were collected. Many factors such as environment conditions, current health conditions of the participants were not disclosed.*__

###(b) Another article titled The School Bully Is Sleepy states the following:62
    "The University of Michigan study, collected survey data from parents on each child's sleep habits
    and asked both parents and teachers to assess behavioral concerns. About a third of the students
    studied were identified by parents or teachers as having problems with disruptive behavior or
    bullying. The researchers found that children who had behavioral issues and those who were
    identified as bullies were twice as likely to have shown symptoms of sleep disorders."
    A friend of yours who read the article says, "The study shows that sleep disorders lead to
    bullying in school children." Is this statement justified? If not, how best can you describe the
    conclusion that can be drawn from this study?

__*Answer: It's difficult to conclude the survey since the response variable or explanatory variable is difficult to derive from the survey. It could be possible that children with sleep disorders behave in a bullying manner, or children with bullying behaviour can show sleep disorder symptoms.*__

### Question  1.36 Exercise and mental health. A researcher is interested in the effects of exercise on
mental health and he proposes the following study: Use stratified random sampling to ensure representative
proportions of 18-30, 31-40 and 41- 55 year olds from the population. Next, randomly
assign half the subjects from each age group to exercise twice a week, and instruct the rest not
to exercise. Conduct a mental health exam at the beginning and at the end of the study, and
compare the results.


(a) What type of study is this? __*Answer: Experimental prospective study using stratified random sampling*__ </br>
(b) What are the treatment and control groups in this study? __*Answer: The group doing exercise is treatment group and the other is control group.*__</br>
(c) Does this study make use of blocking? If so, what is the blocking variable? __*Answer: No, there is no blocking variable mentioned.*__</br>
(d) Does this study make use of blinding? __*Answer: No, there is no blinding in place to create a placebo effect.*__</br>
(e) Comment on whether or not the results of the study can be used to establish a causal relationship
between exercise and mental health, and indicate whether or not the conclusions can
be generalized to the population at large. </br>__*Answer: Since this is an experiment,  a causal relationship can be established and can be applied to population at large.*__
(f) Suppose you are given the task of determining if this proposed study should get funding.
Would you have any reservations about the study proposal?  __*Answer: I would suggest the experiment be conducted under certain conditions to remove blocking variables such as family or work related stresses being applied on an individuals, as this can effect their results.*__


### Question  1.48 Stats scores. Below are the final exam scores of twenty introductory statistics students. 
57, 66, 69, 71, 72, 73, 74, 77, 78, 78, 79, 79, 81, 81, 82, 83, 83, 88, 89, 94 
Create a box plot of the distribution of these scores. The five number summary provided below may be useful.
Min Q1 Q2 (Median) Q3 Max
57 72.5 78.5 82.5 94

```{r}
scores <- c(57, 66, 69, 71, 72, 73, 74, 77, 78, 78, 79, 79, 81, 81, 82, 83, 83, 88, 89, 94)
boxplot(scores)
```

### Question  1.50 Mix-and-match. Describe the distribution in the histograms below and match them to the box plots.

  1.  Histogram 'a' represents a symmetrical unimodal distribution represented by box plot 2
  2.  Histogram 'b' represents a multimodal distribution represented by box plot 3
  3.  Histogram 'c' represents a right skewed distribution represented by box plot 1



### Question  1.56 Distributions and appropriate statistics, Part II . For each of the following, state
whether you expect the distribution to be symmetric, right skewed, or left skewed. Also specify
whether the mean or median would best represent a typical observation in the data, and whether
the variability of observations would be best represented using the standard deviation or IQR.
Explain your reasoning.


(a) Housing prices in a country where 25% of the houses cost below $350,000, 50% of the houses
cost below $450,000, 75% of the houses cost below $1,000,000 and there are a meaningful
number of houses that cost more than $6,000,000. </br>
__*Answer: It is left Skewed since the number of houses higher than 1M is a meaningful number, and makes it a sparsely distributed data on the left. Median and IQR are more suitable to this data due to the skewed nature of data.*__


(b) Housing prices in a country where 25% of the houses cost below $300,000, 50% of the houses
cost below $600,000, 75% of the houses cost below $900,000 and very few houses that cost
more than $1,200,000.
__*Answer: The data is more symmetric in nature due to the equal distribution of housing prices and number of houses in these ranges. Mean and Standard deviation are better suited for this kind of distribution.*__

(c) Number of alcoholic drinks consumed by college students in a given week. Assume that most of
these students don't drink since they are under 21 years old, and only a few drink excessively.</br>
__*Answer This data will be right skewed since majority of students would have a zero drink, with very taking high number of drinks. And this makes the data totally right skewed. Median and IQR are better suited to represent this data.*__

(d) Annual salaries of the employees at a Fortune 500 company where only a few high level
executives earn much higher salaries than the all other employees.</br>
__*Answer This data will be right skewed due to less number of emplyess with high salaries.*__


### Question  1.70 Heart transplants. The Stanford University Heart Transplant Study was conducted to
determine whether an experimental heart transplant program increased lifespan. Each patient
entering the program was designated an official heart transplant candidate, meaning that he was
gravely ill and would most likely benefit from a new heart. Some patients got a transplant and
some did not. The variable transplant indicates which group the patients were in; patients in the
treatment group got a transplant and those in the control group did not. Another variable called
survived was used to indicate whether or not the patient was alive at the end of the study

(a) Based on the mosaic plot, is survival independent of whether or not the patient got a transplant?
Explain your reasoning. </br>
__*Answer: Based on mosaic plot the survival rate of the patients is high for those who got a transplant.*__

(b) What do the box plots below suggest about the efficacy (e???ectiveness) of the heart transplant
treatment.</br>
__*Answer: The survival time(days) is much higher for patients who underwent a transplantation compared to who didn't.*__

(c) What proportion of patients in the treatment group and what proportion of patients in the
control group died?</br>
```{r}
survived_controlGroup <- 4/34
survived_treatmentGroup <- 24/69
```
__*Control group survival proportion is `survived_controlGroup` and treatment group survival proportion is  `survived_treatmentGroup`*__


(d) One approach for investigating whether or not the treatment is effective is to use a randomization
technique.

  i. What are the claims being tested</br>
      __*Answer: Claims about the survival after a heart transplantion*__
  ii. The paragraph below describes the set up for such approach, if we were to do it without
using statistical software. Fill in the blanks with a number or phrase, whichever is
appropriate. </br>


__*We write alive on 4 cards representing patients who were alive at
the end of the study, and dead on 30 cards representing patients
who were not. Then, we shuffle these cards and split them into two groups:
one group of size 34 representing treatment, and another group of
size 69 representing control. We calculate the difference between
the proportion of dead cards in the treatment and control groups (treatment -
control) and record this value. We repeat this 100 times to build a distribution
centered at 0. Lastly, we calculate the fraction of simulations where
the simulated di???erences in proportions are . If this fraction is low,
we conclude that it is unlikely to have observed such an outcome by chance and
that the null hypothesis should be rejected in favor of the alternative.*__

iii. What do the simulation results shown below suggest about the e???ectiveness of the transplant
program?</br>
__*It's a symmetric distribution.*__






    
