---
title: "Assignment_607_Week3"
author: "SantoshCheruku"
date: "February 16, 2019"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
```

##Create Raw data
```{r warning=FALSE}
#Raw Data
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

# Extract only full names and unlist them into a character
names <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}")) 
names

```

## Questions {.tabset}
### **1. First and Last name**
```{r}

# Replace Dr., Rev., C. and any commas to get only first and last names
str_names_only <- str_replace_all(names, "Dr.\\s+|Rev.\\s+|C.\\s+|,", "")

# Split the String into a list of full names
str_names_list <- str_split(noquote(str_names_only), "\\W+")
noquote(str_names_list)
```

### **2. Has a title (i.e., Rev. and Dr.)**
```{r}

str_names_titles <- str_replace_all(names, "C.\\s+|,", "")
logical_vector_title <- c()

str_names_titles <- str_split(str_names_titles, '\\"')
for(full_title_name in str_names_titles) {
  logical_vector_title <- c(logical_vector_title, str_detect(full_title_name, "Dr.|Rev."))
}

logical_vector_title

```

###**3. Has a second name**
```{r}

str_names_second <- str_replace_all(names, "Dr.|Rev.|,", "")
logical_vector_secondName<- c()

str_names_second <- str_split(str_names_second, '\\"')
for(full_title_name in str_names_second) {
  logical_vector_secondName <- c(logical_vector_secondName, str_detect(full_title_name, "\\."))
}

logical_vector_secondName
```

### **4. First and Last name vectors (Extra)**
```{r}

# Replace Dr., Rev., C. and any commas to get only first and last names
str_names_only <- str_replace_all(names, "Dr.\\s+|Rev.\\s+|C.\\s+|,", "")

# Split the String into a list of full names
str_names_list <- str_split(noquote(str_names_only), "\\W+")
str_names_list
#Create two empty vectors to hold first and last names
first_names_vector <- c()
last_names_vector <- c()

#Use a for loop to iterate over the List of full names and extract first and last names into vectors
for (full_name in str_names_list) { 
    first_names_vector <- c(first_names_vector, full_name[1])  
    last_names_vector <- c(last_names_vector, full_name[2])
}
first_names_vector
last_names_vector
```

##Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.  {.tabset}
### **1. [0-9]+\\$ **
Below regular expression will match numbers ending with a $ sign
```{r}
sample_string <- "220$ is the price of this jacket bought on date 10/12/2018"
str_extract_all(sample_string, "[0-9]+\\$")
```

### **2  \\b[a-z]{1,4}\\b**
Below regular expression will match words starting with small alphabets with a count of letters b/w 1 to 4 ending with a $ sign
```{r}
sample_string <- "220$ is the price of this jacket bought on date 10/12/2018."
str_extract_all(sample_string, "\\b[a-z]{1,4}\\b")
```
### 3 .*?\\.txt$
```{r}
# Below regular expression will match words having any or no characters and ending with a .txt
sample_text_names <- c('filename', 'file.txt', 'file.xml')
str_extract_all(sample_text_names, ".*?\\.txt$")
```
### **4 \\d{2}/\\d{2}/\\d{4}**
Below regular expression will match numbers in the format of dd/dd/dddd usually a date format 
```{r}
sample_string <- "220$ is the price of this jacket bought on date 10/12/2018"
str_extract_all(sample_string, " \\d{2}/\\d{2}/\\d{4}")
```
### **5 <(.+?)>.+?</\\1> **
Below Regular expression will match a literal '<' followed by any chars until > having an optional chars before matching '</' followed a match which is a back reference to first match. Usualy used to extract xml tags and values
```{r}
sample_string <- "This text having <sometag>value</sometag> is a sample xml tag"
str_extract_all(sample_string, "<(.+?)>.+?</\\1>")
```


##Decrypting String
When tested using upper predefined character class following message reveals 'CONGRATULATIONSYOURARESUPERNERD'
```{r}
encrypted <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

unencrypted <- unlist(str_extract_all(encrypted, "[:upper:]"))
unencrypted
```