---
title: "601_HW1_HTML"
author: "SantoshCheruku"
date: "9/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(corrplot)
```

## Load datasets
```{r}
train <- read.csv("datasets/moneyball-training-data.csv")
train <- train[c(2:17)]
summary(train)

head(train, n=10)

```

## Exploratory Data Analysis


### Wins by Team
```{r}
plot(train$TEAM_BATTING_H, train$TARGET_WINS )
```


```{r}
par(mfrow=c(2,2))
hist(train$TEAM_BATTING_H,
     main = "hits histogram", xlab = "hits (season)",
     breaks = 20)
hist(train$TEAM_BATTING_2B,
     main = "doubles histogram", xlab = "doubles (season)",
     breaks = 20)
hist(train$TEAM_BATTING_3B,
     main = "triples histogram", xlab = "triples (season)",
     breaks = 20)
hist(train$TEAM_BATTING_HR,
     main = "homeruns histogram", xlab = "homeruns (season)",
     breaks = 20)
par(mfrow=c(1,1))
```


### Addressing data quality issues

From the summary, we can observe that following variables have a high number of NA or missing values, and we need to address these before we can work on the models.

TEAM_BATTING_SO  - 102</br>
TEAM_BASERUN_SB  - 131</br>
TEAM_BASERUN_CS - 772</br>
TEAM_BATTING_HBP - 2085</br>
TEAM_PITCHING_SO - 102</br>
TEAM_FIELDING_DP - 286</br>

TEAM_BATTING_HBP has highest number of missing cases i.e., 2085 (almost 90%). Therefore, we can drop this variable from our dataset.
```{r}
# Removing TEAM_BATTING_HBP variable
 train <- subset(train, select = -c(10))
```

Now let's measure how many complete observations do we have

```{r}
sum(complete.cases(train)==TRUE)

sum(complete.cases(train)==FALSE)
```

## Set mean value to the missing col values in the dataset
```{r}

train$TEAM_BATTING_SO[is.na(train$TEAM_BATTING_SO)] = mean(train$TEAM_BATTING_SO, na.rm = TRUE)
train$TEAM_BASERUN_SB[is.na(train$TEAM_BASERUN_SB)] = mean(train$TEAM_BASERUN_SB, na.rm = TRUE)
train$TEAM_BASERUN_CS[is.na(train$TEAM_BASERUN_CS)] = mean(train$TEAM_BASERUN_CS, na.rm = TRUE)
train$TEAM_PITCHING_SO[is.na(train$TEAM_PITCHING_SO)] = mean(train$TEAM_PITCHING_SO, na.rm = TRUE)
train$TEAM_FIELDING_DP[is.na(train$TEAM_FIELDING_DP)] = mean(train$TEAM_FIELDING_DP, na.rm = TRUE)


```

We have about 1486 complete observations, which is close to 65%. Therefore, we should try to work with other variables to see if we can increase this as much as possible.


We have 2 options, either replace them with a mean or alternate value or identify alias variables which represent them with similar correlation.


### Identify highest impacting independent variables
```{r}
lm <- lm(formula = train$TARGET_WINS ~ ., data = train)
summary(lm)
```


### Use covariates
```{r}
#cor(train)

cor_train <- cor(train,  use = "na.or.complete")
corrplot(cor_train)
```

