---
title: "GradStudent_STEM_Impact_606FinalProject"
author: "SantoshCheruku"
date: "April 28, 2019"
output:
  html_document:
    toc: true
    toc_float: true

---

```{css, echo=FALSE}
.rCode {
background-color: black;
color: white;
}
```

```{r, warning=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

#Introduction
Choice of major in graduation has always been a major decision in every student's life. Having a good estimate about the job and salary prospects in a major in future defintely helps students in making an informative decision about his/her grad choice. 
We have seen numerous articles advising which course to choose and which one would be in demand with prospective employers. 'STEM' which stands for **S**cience **T**echnology **E**ngineering **M**athematics, and has been highlighted to be a great choice in majors to find high paying jobs. </br>
**In this project, I'm researching to find whether the choice of STEM major in graduation has an impact on a graduate salary? **

#Data

**Data collection:**The recent grad list compiled by census.gov has provided us with a list of compiled data of grads during 2010-2012. There are observations made against 173 majors. </br>
**Cases:** Each observation represents a major with data mentioning the number of grads who passed, their employement rate, salary ranges etc. There are totally 173 observations which each observation representing a major.</br>
**Variables:**We will be studying two variables, the median salary which is a response variable and course major which is an independent variable also an explantory variable. This is an observation study since it's derived from a data which is taken from an observation during a period of time. </br>
**The population of interest** is the graduates of United States. And the findings from this analysis can be generalized for the entire population of interest i.e., the graduates. </br>
**Causality:** We can use this data to establish a causality between the variables of interest. Since the data has been sampled randomly from a large population of data source.

Source: https://github.com/fivethirtyeight/data/tree/master/college-majors


```{r block0, class.source="rCode"}
grads <- read.csv("https://raw.githubusercontent.com/san123i/CUNY/master/Semester1/606/Final%20Project/grad-students.csv")
head(grads)%>%
  kable() %>%
  kable_styling() %>% scroll_box(width = "800px")


summary(grads)
```


#Exploratory Data Analysis

##1. Number of graduates in each major
Let's do an analysis on the number of graduates in each major. </br>

```{r block1, class.source="rCode"}
isSTEM <- function(majorCategory){
    result <- grepl("science|math|engineering|technology|computer", majorCategory, ignore.case = T)
    result <- ifelse(result, !grepl("Social", majorCategory, ignore.case = T), result)
    return(ifelse(result, "STEM","NON-STEM"));
}
```
The below plot shows that majority of the graduates fall under 'Education' major type where as the least are under 'interdiscriplinary' and 'Agriculture & Natural Resources' category. Also, there are only 4 major categories which fall under STEM i.e. 'Biology & Life Sciences', 'Computers & Mathematics', 'Engineering' and 'Physical Sciences'.

```{r, class.source="rCode"}
grads <- grads %>% mutate("Major_Type_STEM"=isSTEM(Major_category))

x <- grads %>% select(Major, Major_category, Major_Type_STEM)
#x[order(x$Major_Type_STEM),]
x[order(x$Major_Type_STEM),] %>%
  kable() %>%
  kable_styling()

ggplot(grads, aes(x=reorder(Major_category, Grad_total), y = Grad_total, fill=Major_Type_STEM)) + geom_bar(stat="identity") + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))+ labs( title = 'Graduates by major category type')

list <- aggregate(Grad_total ~ Major_category, grads, sum)
ggplot(list, aes(x=reorder(Major_category, Grad_total), y=Grad_total, label=Grad_total, label.size=.25)) + geom_col() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))+ labs(title = '# of Graduates by major category type') + 
  geom_text(size = 3, position = position_stack(vjust = 0.5), color='white')

list <- aggregate(Grad_total ~ Major_Type_STEM, grads, sum)
ggplot(list, aes(x=reorder(Major_Type_STEM, Grad_total), y=Grad_total, label=Grad_total, fill=Major_Type_STEM)) + geom_col() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))+ labs(title = '# of Graduates by STEM type') + 
  geom_text(size = 3, position = position_stack(vjust = 0.5), color='white')

# label=sum(Grad_total)
```

##2. Salary data analysis by category

The below plot shows the grad median salary in each major category. 

```{r, class.source="rCode"}  

ggplot(grads, aes(Major_category, Grad_median, fill=Major_Type_STEM)) + geom_boxplot()  +  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

ggplot(grads, aes(x=Major, y=Grad_median, shape=Major_Type_STEM, colour=Major_Type_STEM)) + geom_point()  + scale_shape_manual(values=c(19, 2))
```


##3. Average 'p25', 'p75' and 'Median' salary of each category.
```{r, class.source="rCode"}  
ggplot(grads, aes(x=reorder(Major_category, Grad_P25), y = Grad_P25, fill=Major_Type_STEM)) + geom_bar(stat="identity") +labs(title = 'a. 25Perc of Grad salary by Major_category') +  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

ggplot(grads, aes(x=reorder(Major_category, Grad_P75), y = Grad_P75, fill=Major_Type_STEM)) + geom_bar(stat="identity") +labs(title = 'b. 75Perc of Grad salary by Major_category') +  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

ggplot(grads, aes(x=reorder(Major_category, Grad_median), y = Grad_median, fill=Major_Type_STEM)) + geom_bar(stat="identity") +labs(title = 'c. Median of Grad salary by Major_category') +  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

##4. Grad Median salaries by STEM or Non-STEM
```{r, class.source="rCode"}  
ggplot(grads, aes(Major_Type_STEM, Grad_median, fill=Major_Type_STEM)) + geom_boxplot()  +  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

```

#Inference
We are using Analysis of Variance (ANOVA) method to dervie an inference from the Hypothesis. Below are the conditions we evaluated and found to be satisfied. </br>
a. Independence of cases - The cases provided as per data are independent in nature </br>
b. Distribution of Residuals are Normal -  The below qqnorm graph displays that residuals are normally distributed without major outliers</br>
c. Homogentiy of variances</br>

Let's start our inference with a Null Hypothesis test</br>
**H0 <- The median salary for an average graduate in STEM and Non-STEM categories are same </br>**
**HA <- The median salary for an average graduate in STEM and non-STEM categories are not same.</br>**

```{r}
grads_anova <- aov(grads$Grad_median ~ grads$Major_Type_STEM)
summary(grads_anova)

plot(grads_anova)
```

Since the P value is less than .05, we reject the null hypothesis (H0), and can conclude that average graduate salary for STEM and non-STEM category students is not equal. And from the geom plot, we can identify that Median salary from STEM category students is significantly higher than non-STEM categories.


#Conclusion
Using the given data, we were able to do an exploratory analysis and identify a clear pattern/trend that STEM course major students have a median salary higher than non-grad students. And using the ANOVA analysis, we were able to confirm that grad median salaries for both STEM and Non-STEM grad student are NOT equals. Therefore, we can clearly state that grad median salaries are higher for STEM students. </br>
**Future research:** I believe there are few other factors which can influence the grad salary such as market conditions, number of open job opportunities, type of economies i.e., a developed, emerging or under developed. If I can get access to these datasets, then an even more realistic and robust model can be designed. 